\chapter{Solutions to Exercises}
\footnotesize

\section*{Chapter \ref{chap:fundamentals}}

\begin{sltn}[\ref{exer:setops}]~
\begin{enumerate}[label=(\alph*)]
    \item $ A\cup B=\{-1,1,2,3\} $,
    
    $ A\cap B=\{1\} $,

    $ A\times B=\{(-1,1),(-1,2),(-1,3),(1,1),(1,2),(1,3)\} $,

    $ B\times A=\{(1,-1),(1,1),(2,-1),(2,1),(3,-1),(3,1)\} $.

    \item $ A\cup B=\{0,1,3,9,27\} $,
    
    $ A\cap B=\varnothing $,

    $ A\times B=\{(0,3),(0,9),(0,27),(1,3),(1,9),(1,27)\} $,

    $ B\times A=\{(3,0),(3,1),(9,0),(9,1),(27,0),(27,1)\} $.

    \item $ A\cup B=[-1,\infty) $,
    
    $ A\cap B=(0,1] $,

    $ A\times B=\{(a,b)\mid a\in[-1,1],b\in(0,\infty)\} $,

    $ B\times A=\{(b,a)\mid b\in(0,\infty),a\in[-1,1]\} $.\qedhere
\end{enumerate}
\end{sltn}

\begin{sltn}[\ref{exer:Z}]
$ \mathbb{Z}=\{\pm n\mid n\in\mathbb{N}\}\cup\{0\} $.
\end{sltn}

\begin{sltn}[\ref{exer:C}]
$ \mathbb{C}=\{a+bi\mid a,b\in\mathbb{R}\} $.
\end{sltn}

\section*{Chapter \ref{chap:vectors}}

\begin{sltn}[\ref{exer:cfield}]
Let $ a,b,c,d,e,f\in\mathbb{R} $. First we must verify that $ \mathbb{C} $ is indeed closed under the given operations. Since $ (a+c),(b+d)\in\mathbb{R} $, we see that
\begin{equation*}
    (a+bi)+(c+di)=(a+c)+(b+d)i\in\mathbb{C},
\end{equation*}
and similarly, since $ (ac-bd),(ad+bc)\in\mathbb{R} $, we also see that
\begin{equation*}
    (a+bi)\cdot(c+di)=(ac-bd)+(ad+bc)i\in\mathbb{C}.
\end{equation*}
Thus, $ \mathbb{C} $ is closed under $ + $ and $ \cdot $. Now we can check the field axioms:
\begin{enumerate}
    \item Associativity:
    \begin{align*}
        ((a+bi)+(c+di))+(e+fi) &= (a+c+(b+d)i)+(e+fi) \\
        &= a+c+e+(b+d+f)i \\
        &= a+bi+(c+e)+(d+f)i \\
        &= (a+bi)+((c+di)+(e+fi)).
    \end{align*}
    \begin{align*}
        ((a+bi)\cdot(c+di))\cdot(e+fi) &= (ac-bd+(ad+bc)i)\cdot(e+fi) \\
        &= (ac-bd)e-(ad+bc)f+((ac-bd)f+(ad+bc)e)i \\
        &= ace-bde-adf-bcf+(acf-bdf+ade+bce)i \\
        &= a(ce-df)-b(cf+de)+(a(cf+de)+b(ce-df))i \\
        &= (a+bi)\cdot(ce-df+(cf+de)i) \\
        &= (a+bi)\cdot((c+di)\cdot(e+fi)).
    \end{align*}

    \item Commutativity:
    \begin{align*}
        (a+bi)+(c+di) &= (a+b)+(c+d)i=(b+a)+(d+c)i=(c+di)+(a+bi).
    \end{align*}
    \begin{align*}
        (a+bi)\cdot(c+di) &= (ac-bd)+(ad+bc)i=(ca-db)+(da+cb)i=(c+di)\cdot(a+bi).
    \end{align*}

    \item Existence of additive identity: Consider $ 0=0+0i\in\mathbb{C} $. We see
    \begin{equation*}
        (a+bi)+(0+0i)=(a+0)+(b+0)i=a+bi.
    \end{equation*}

    \item Existence of multiplicative identity: Consider $ 1=1+0i\in\mathbb{C} $. We see
    \begin{equation*}
        (a+bi)\cdot(1+0i)=(1a-0b)+(0a+1b)i=a+bi.
    \end{equation*}

    \item Existence of additive inverse: For $ a+bi $, consider $ -a-bi\in\mathbb{C} $. We see
    \begin{equation*}
        (a+bi)+(-a-bi)=(a-a)+(b-b)i=0+0i=0.
    \end{equation*}

    \item Existence of multiplicative inverse: Suppose $ a+bi\neq 0 $. Consider\footnote{This can be found by setting $ (a+bi)\cdot(x+y i)=1 $ and solving for $ x $ and $ y $.}
    \begin{equation*}
        \frac{a}{a^2+b^2}-\frac{b}{a^2+b^2}i\in\mathbb{C}.
    \end{equation*}
    We see
    \begin{align*}
        (a+bi)\cdot\left(\frac{a}{a^2+b^2}-\frac{b}{a^2+b^2}i\right) &= \frac{a^2}{a^2+b^2}+\frac{b^2}{a^2+b^2}+\left(\frac{ab}{a^2+b^2}-\frac{ba}{a^2+b^2}\right)i \\
        &= \frac{a^2+b^2}{a^2+b^2}+0i=1.
    \end{align*}

    \item Distributivity:
    \begin{align*}
        (a+bi)\cdot((c+di)+(e+fi)) &= (a+bi)\cdot(c+e+(d+f)i) \\
        &= a(c+e)-b(d+f)+(a(d+f)+b(c+e))i \\
        &= ac+ae-bd-bf+(ad+af+bc+be)i \\
        &= ac-bd+(ad+bc)i+ae-bf+(af+be)i \\
        &= ((a+bi)\cdot(c+di))+((a+bi)\cdot(e+fi)).
    \end{align*}
\end{enumerate}
Hence, $ \mathbb{C} $ with these operations is a field.
\end{sltn}

\begin{sltn}[\ref{exer:0unique}]
By the definition of the zero vector, for any vector $ \vect{v} $, we have $ \vect{v}+\vect{0}=\vect{v} $ and $ \vect{v}+\tilde{\vect{0}}=\vect{v} $. In particular, $ \vect{0}=\vect{0}+\tilde{\vect{0}}=\tilde{\vect{0}}+\vect{0}=\tilde{\vect{0}} $.
\end{sltn}

\begin{sltn}[\ref{exer:proof-0subspace} (Proof of Lemma \ref{lem:0subspace})]
Since $ W $ is a vector space, there must exist a zero vector $ \vect{0}_W\in W $, and since $ W\subseteq V $, we have $ \vect{0}_W\in V $. By uniqueness of the zero vector (see Exercise \ref{exer:0unique}), $ \vect{0}_W=\vect{0}_V $.
\end{sltn}

\begin{sltn}[\ref{exer:proof-subspaceintersect-n} (Proof of Corollary \ref{cor:subspaceintersect-n})]
We can prove this result by induction. The base case (where $ n=2 $) is Theorem \ref{thm:subspaceintersect-2}. For our inductive hypothesis, suppose $ U=\cap_{k=1}^m W_k $ is a subspace of $ V $. By Theorem \ref{thm:subspaceintersect-2}, if $ W_{m+1} $ is a subspace of $ V $, then $ U\cap W_{m+1}=\cap_{k=1}^{m+1}W_k $ is also a subspace of $ V $. Hence, $ \cap_{k=1}^n W_k $ is a subspace of $ V $ for all $ n\geq 2 $.
\end{sltn}